Simulation of Nano Particles in a Laser Trap

### 2017-01-26 doc/draft2.tex
updates

\begin{document}
\title{Temperature Measurements in Optical Tweezer Experiments}
\author{Mathias H\"old, BSc.}
\date{2016}
\maketitle
\thispagestyle{empty}
\newpage
\section{Introduction}
%something about computer science in general, optical tweezers and the need to study the
%systems on computers yada yada

@@

\section{The experiment}

@@

laser beam acting on the particle. The experimental setup is depicted in fig.~\ref{fig:setup}.\\

@@

@@

@@

The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in fig. \ref{fig:fcc}.

@@

consisting of four atoms, as shown in fig. \ref{fig:unitcell}, which can be described by a set of points

@@

t$ and the equations of motion are solved for every segment separately (see fig. \ref{fig:finitedifference}).\\

@@

Hamiltonian regions (see fig. \ref{fig:ehex}). 

@@

notation of fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 

@@

\newpage
\section{Results}
% results with different parameters, graphs, etc.
Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
\subsection{The Crystal}
The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
can be seen in fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
where the atoms take the position on the gap between the atoms from the layer below.\\
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
        \caption{A rendered image of the FCC lattice. The chosen color-coding along the (1,1,1) Miller-index 
                highlights the layers for this perspective. The perspective is
                chosen to see the layering of the FCC lattice with its 3 layer structure. The black box acts as a reference point to outline the 
                geometry of the system.}
            %Rendering of the model of the nano crystal before starting the simulation. 
            %The chosen perspective shows the 3 layers of the FCC crystal with a
                    %These layers consist of atoms
                    %taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
                %below. The black box is the outline of the particles' positions and shows the geometry of the system.}
        \label{fig:glass}
    \end{center}
\end{figure}

@@

that it contradicts the constant energy property of the ensemble. If we look at the energy values in fig.~\ref{fig:energystart} we can see the

@@

\subsection{eHEX}
The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
algorithm. In fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_te.pdf}
        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
            and the values for the energy on the right hand side. }
        \label{fig:ehex_te}
    \end{center}
\end{figure}
The comparison of different values of $\Delta Q$ can be seen in fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
been shown in fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
        \label{fig:ehex_dq}
    \end{center}
\end{figure}

@@

all three of these stages are represented by different gas particles, as can be seen in fig.~\ref{fig:barointeraction}.\\

>>>>

\begin{document}
\title{Simulation of Nano Particles in a Laser Trap}
\author{Mathias H\"old, BSc.}
\date{2016}
\maketitle
\thispagestyle{empty}
\newpage
\section{Introduction}
%something about computer science in general, optical tweezers and the need to study the
%systems on computers yada yada

@@

@@

laser beam acting on the particle. The experimental setup is depicted in Fig.~\ref{fig:setup}.\\

@@

    \label{eq:langevin}

@@

\subsection{Model of the Surrounding Gas}
J. Millen et al.~\cite{MillenJ.2014} worked on a similar experimental setup (without the feedback mechanism) and investigated the heating of the
particle in the trap as well as the gas surrounding the particle.\\
They start from the premise that there are 4 different temperatures one has to consider in such an experiment: the temperature of the gas particles
before interacting with the nano particle $T_\text{imp}$ (for \textit{impinging}), the temperature of the gas particles after 
interaction $T_\text{em}$ (for emerging), the surface temperature of the nano particle $T_\text{sur}$ and the temperature of the 
center of mass of the nano particle $T_\text{COM}$. The experimental setup
and the different temperatures are depicted in Fig.~\ref{fig:levitation}.\\
        \includegraphics[scale=0.2]{images/nnano_millen.jpg}
        \caption{The experimental setup used by Millen et al. In \textbf{a} the experiment is schematically depicted and \textbf{b} shows a photograph
            of the actual experimental setup. In \textbf{c} the 4 different temperatures that are present during the experiment are depicted and
        \textbf{d} shows the position of the particle in the trap over time. The image was taken from~\cite{MillenJ.2014}}
        \label{fig:levitation}
Since the two temperatures of the gas, $T_\text{imp}$ and $T_\text{em}$, are not equal, the gas surrounding the nano particle is not in a thermal
equilibrium, which is in contrast to the assumption in many optical tweezer experiments. Since the gas particles only interact with the nano particle
in the trap and not wich each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator
between those two baths. This situation is depicted in
Fig.~\ref{fig:heatbaths}.
        \includegraphics[scale=0.5]{images/nano_nonequilibrium_cropped.jpg}
        \caption{Schematic picture of the two heat baths. The (hot) nano particle interacts with the (cold) incoming gas particles and they leave with
        a higher temperature. The incoming and outgoing particles do not interact with each other and thus create two separate heat baths with
    different temperatures. Image modified from \cite{Kroy2014}}
        \label{fig:heatbaths}
The one dimensional Langevin equation \eqref{eq:langevin} has to be modeled to include the different friction coefficients of the two heat baths. The
model derived in \cite{MillenJ.2014} is described in the following equation:
    M\ddot{x}(t) + M\left(\Gamma^\text{imp}+\Gamma^\text{em}\right)\dot{x}(t) + M\omega^2x(t) =F^\text{imp}+F^\text{em}
where $x$ is the position of the nano particle, $\dot{x}$ and $\ddot{x}$ are the velocity and the acceleration, respectively, $M$ is the mass,
$\omega$ is the trap frequency, $\Gamma^\text{imp}$ and $\Gamma^\text{em}$ are the damping constants for the 
impinging particles and the emerging particles, respectively and $F^\text{imp}$ and $F^\text{em}$ are the associated noise terms, 
which satisfy $\langle F^\text{imp}\rangle= \langle F^\text{em} \rangle = 0$ individually.\\
Miller et al. investigate the dependence of the emerging temperature $T_\text{em}$ on the intensity of the laser for different sized nano particles
and different pressures.\\

%These four temperatures introduced in the paper, together with the surrounding pressure and the intensity of the laser 
%The existence of the two temperatures of the gas implies that, contrary to the assumption in many of the experiments, the gas is not in a thermal
%equilibrium. The Langevin equation (like the one in \eqref{eq:langevin}) has to be modeled k

@@

The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in Fig. \ref{fig:fcc}.

@@

consisting of four atoms, as shown in Fig. \ref{fig:unitcell}, which can be described by a set of points

@@

t$ and the equations of motion are solved for every segment separately (see Fig. \ref{fig:finitedifference}).\\

@@

Hamiltonian regions (see Fig. \ref{fig:ehex}). 

@@

notation of Fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 

@@

\newpage
\section{Results}
% results with different parameters, graphs, etc.
Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
\subsection{The Crystal}
The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
can be seen in Fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
where the atoms take the position on the gap between the atoms from the layer below.\\
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
        \caption{A rendered image of the FCC lattice. The chosen color-coding along the (1,1,1) Miller-index 
                highlights the layers for this perspective. The perspective is
                chosen to see the layering of the FCC lattice with its 3 layer structure. The black box acts as a reference point to outline the 
                geometry of the system.}
            %Rendering of the model of the nano crystal before starting the simulation. 
            %The chosen perspective shows the 3 layers of the FCC crystal with a
                    %These layers consist of atoms
                    %taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
                %below. The black box is the outline of the particles' positions and shows the geometry of the system.}
        \label{fig:glass}
    \end{center}
\end{figure}

@@

that it contradicts the constant energy property of the ensemble. If we look at the energy values in Fig.~\ref{fig:energystart} we can see the

@@

\subsection{eHEX}
The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
algorithm. In Fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_te.pdf}
        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
            and the values for the energy on the right hand side. }
        \label{fig:ehex_te}
    \end{center}
\end{figure}
The comparison of different values of $\Delta Q$ can be seen in Fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
been shown in Fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
        \label{fig:ehex_dq}
    \end{center}
\end{figure}

@@

all three of these stages are represented by different gas particles, as can be seen in Fig.~\ref{fig:barointeraction}.\\
